### 2015-05-16 ase-2015-tex/summary-heap.tex

\input{surface-syntax}
\input{machine-syntax}
\begin{figure}[t]
\begin{center}
\begin{tabular}{c}
\scalebox{0.9}{\usebox{\boxSurface}} \\
(a) \\\\
\scalebox{0.9}{\usebox{\boxMachine}} \\
(b)
\end{tabular}
\end{center}
\caption{Syntax for Javalite. (a) The Javalite surface syntax. (b) The machine syntax with $\bowtie\ \in \{\wedge,\vee,\Rightarrow\}$.}
\label{fig:syntax}
\end{figure}

A program, $\cfgnt{P}$, is a registry of classes, $\mu$, with
a tuple indicating a class, $\cfgnt{C}$, and a method, $\cfgnt{m}$,
where execution starts. The only primitive type in Javalite is
Boolean. For simplicity in presentation, classes have only
non-primitive fields, and methods have a single
parameter. Expressions, $\cfgnt{e}$, include statements, and they use
'$\cfgt{:=}$' to indicate assignment and '$\cfgt{=}$' to indicate
comparison.  The dot-operator for field access is replaced by
'$\cfgt{\$}$', and the dot-operator for method invocation is replaced
by '$\cfgt{@}$' since '.' is reserved in PLT Redex. There is no
explicit return statement; rather, the value of the last expression is
used as the return value. A variable is always indicated by \cfgnt{x}
and a value by \cfgnt{v}. A value can be a reference in the heap,
$\cfgnt{r}$, or any of the special values shown
in~\figref{fig:syntax}(a).  Only type correct programs are given as
input to the machine.

The state of the machine, $\cfgnt{s}$, includes the program
registry, $\mu$, a pair of functions defining the heap
$\lp\cfgnt{L}\ \cfgnt{R}\rp$, the path condition, $\phi$, a constraint defined over
references in the heap, the environment, $\eta$, for local variables,
and the continuation $\cfgnt{k}$. The registry never changes so it is
omitted from the state tuple in the rest of the presentation.  The
more complex structures, such as the environment and heap, are defined
as lists that start with empty, \cfgnt{mt}. These lists are treated as
partial functions so $\eta(\cfgnt{x}) = \cfgnt{r}$ is the reference
$\cfgnt{r}$ associated with variable $\cfgnt{x}$. The notation
$\eta^\prime = \eta[\cfgnt{x} \mapsto \cfgnt{v}]$ defines a new
partial function $\eta^\prime$ that is the same as $\eta$ except that
the variable $\cfgnt{x}$ now maps to $\cfgnt{v}$. The continuation
$\cfgnt{k}$ indicates where the hole, $\cfgt{*}$, is located, stores
temporary computation, and keeps track of the next computation. For
example, the continuation $\lp\cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$ indicates that the machine is evaluating the
expression for the object reference on which the field $\cfgnt{f}$ is
going to be accessed. Once the field access is complete, the machine
continues with $\cfgnt{k}$.

>>>>

\input{surface-syntax}
\input{machine-syntax}
\begin{figure}[t]
\begin{center}
\begin{tabular}{c}
\scalebox{0.9}{\usebox{\boxSurface}} \\
(a) \\\\
\scalebox{0.9}{\usebox{\boxMachine}} \\
(b)
\end{tabular}
\end{center}
\caption{ (a) The surface syntax. (b) The machine syntax with $\bowtie\ \in \{\wedge,\vee,\Rightarrow\}$.}
\label{fig:syntax}
\end{figure}

A program, $\cfgnt{P}$, is a registry of classes, $\mu$, with
a tuple indicating a class, $\cfgnt{C}$, and a method, $\cfgnt{m}$,
where execution starts. For simplicity in presentation, booleans are the only primitive type, 
classes have only non-primitive fields, and methods have a single
parameter. Expressions, $\cfgnt{e}$, include statements, and they use
'$\cfgt{:=}$' to indicate assignment and '$\cfgt{=}$' to indicate
comparison.  The dot-operator for field access is replaced by
'$\cfgt{\$}$', and the dot-operator for method invocation is replaced
by '$\cfgt{@}$' since '.' is reserved in PLT Redex. There is no
explicit return statement; rather, the value of the last expression is
used as the return value. A variable is always indicated by \cfgnt{x}
and a value by \cfgnt{v}. A value can be a reference in the heap,
$\cfgnt{r}$, or any of the special values shown
in~\figref{fig:syntax}(a).  Only type correct programs are given as
input to the machine.

The state of the machine, $\cfgnt{s}$, includes the program
registry, $\mu$, a pair of functions defining the heap
$\lp\cfgnt{L}\ \cfgnt{R}\rp$, represented in the syntactic machine as singly linked lists of mappings, 
the path condition, $\phi$, a constraint defined over
references in the heap, the environment, $\eta$, for local variables,
and the continuation $\cfgnt{k}$. The registry never changes so it is
omitted from the state tuple in the rest of the presentation.  The
more complex structures, such as the environment and heap, are defined
as lists that start with empty, \cfgnt{mt}. These lists are treated as
partial functions so $\eta(\cfgnt{x}) = \cfgnt{r}$ is the reference
$\cfgnt{r}$ associated with variable $\cfgnt{x}$. The notation
$\eta^\prime = \eta[\cfgnt{x} \mapsto \cfgnt{v}]$ defines a new
partial function $\eta^\prime$ that is the same as $\eta$ except that
the variable $\cfgnt{x}$ now maps to $\cfgnt{v}$. The continuation
$\cfgnt{k}$ indicates with $\cfgnt{*} where the expression $\cfgnt{e}$ came from, stores
temporary computation, and keeps track of the next continuation. For
example, the continuation $\lp\cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$ indicates that the machine is evaluating the
expression for the object reference on which the field $\cfgnt{f}$ is
going to be accessed. Once the field access is complete, the machine
continues with $\cfgnt{k}$.
